---
title: "Final Project"
author: "F.R.Castilla, O.Contrera"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newpage

# Table of Contents
## 1. [Abstract] 
## 2. [Description Data]
## 3. [Technique]
## 4. [Analysis]
## 5. [Conclusion]
## 6. [Bibliografy]
## 7. [Apendix]

\newpage
## Abstract
Lupus is a chronic, complex autoimmune disease that can affect the joints, skin, brain, lungs, kidneys and blood vessels, causing widespread inflammation and tissue damage in the affected organs.

In order to study whether early detection of this disease is relevant to its survival, we used 87 people followed up for this study.

for this study we used 87 followed for 15+ years there have been 35 deaths, the time the disease has appeared and the duration which is the time it took to be detected. 


## Description data
the data has been extracted from http://lib.stat.cmu.edu/datasets/lupus

TIME(int) time during which the disease has occurred
STATUS(int) 0 alive 1 dead
DURATION(num) time during which the viropsy was taken
LOG.1.DURATION.(num) log(1+duration)



```{r}
data<-data.frame(read.table("~/Statistic/Project/lupus.txt",header = T))
dim(data)
attach(data)
```
first we see that our table has a dimension of 87x4

our objective would be to see if the STATUS which is what tells us if the subject is alive (0) or dead (1).

as a first step we see how many of the subjects in the study have died or are alive.

```{r}
table(STATUS)
```
we can see that at the end of the study there have been a total of 35 deaths out of 87 over the 15+ years of the study. 

## Technique 

as our objective is to be able to see which predictors are the most important for predicting the STATUS and this is a binomial variable (0, 1) the model that can be most closely approximated is the logistic regression model. 

## Analysis

```{r}
mean(data$DURATION[data$STATUS==0])
mean(data$DURATION[data$STATUS==1])
mean(data$DURATION)
```
with a mean of 10.48 being the time the subjects had the disease until they were biopsied and started treatment we can see that the mean for the live subjects is 9.23 and the mean for the deceased is 12.34 and we can begin to see that there is a difference between the time the disease was started and the time it was treated. 

```{r}
boxplot(DURATION~STATUS)
```

in this boxplot it cannot be distinguished well so in order to work with a more comfortable scale we will work with the logarithm+1.


```{r}
boxplot(LOG.1.DURATION.~STATUS)
```
where you can already see some difference but still not enough to be able to see if there is a relationship between the time that passed before the treatment started and whether it survives or not we create our model, also as we are dealing with a data that is binary live/dead I will use a STATUS~DURATION logistic model.

```{r}
model1.1<-glm(STATUS~DURATION,family = binomial(link = 'logit'), trace=TRUE)
summary(model1.1)
anova(model1.1,test="Chisq")
dchisq(0.43916 ,1)
```

with the summary of the model we can already start to see that the DURATION variable does not have an influence. 

```{r}
plot(STATUS~DURATION)+
points(DURATION,fitted(model1.1),pch=1,col="blue")

```
so we are going to look if there is another model that can explain the deaths so we are going to try STATUS~DURATION+TIME
and STATUS~TIME,

```{r}

model1.2<-glm(STATUS~DURATION+TIME,family = binomial(link = 'logit'), trace=TRUE)
summary(model1.2)
anova(model1.2,test="Chisq")
dchisq(36.619 ,1)
dchisq(0.439,1)
```

```{r}
model1.3<-glm(STATUS~TIME,family = binomial(link = 'logit'), trace=TRUE)
summary(model1.3)

anova(model1.3,test="Chisq")
dchisq(36.619 ,1)
```



I think the model that can be 3 is the one that fits best STATUS~TIME


```{r}
plot(STATUS~TIME)+
points(TIME,fitted(model1.3),pch=1,col="blue")
```


Another model to look at is the logarithm of DURATION.

```{r}
model1.4<-glm(STATUS~LOG.1.DURATION.,family = binomial(link = 'logit'), trace=TRUE)
summary(model1.4)

anova(model1.4,test="Chisq")
dchisq(1.044 ,1)
```

```{r}
plot(STATUS~LOG.1.DURATION.)+
points(LOG.1.DURATION.,fitted(model1.4),pch=1,col="blue")
```
In the end the best model looking at the anova test is model 1.3 STATUS~TIME. 


To see how much the model explains the variance we can use the pseudoR, to calculate the pseudoR we first have to make the model null and then use the formula
$R^{2}_{McFadden}=1-\frac{likelihood model}{likelihoodnullmodel}$

```{r}
nullmodel<-glm(STATUS~1,family = binomial(link = 'logit'), trace=TRUE)
PseudoR1 <- 1-logLik(model1.3)/logLik(nullmodel)
PseudoR1
PseudoR <- 1-logLik(model1.1)/logLik(nullmodel)
PseudoR

PseudoR2<- 1-logLik(model1.4)/logLik(nullmodel)
PseudoR2

```
With this we can see that while TIME can explain 32% of the variance, only 0.3% of the variance can be explained by DURATION and with its logarithm of 0.8% an improvement can be seen but it still does not explain the variance.

At the end we calculate the coefidence interval of the models using the formula
$CI=\bar{x}\pm z .*SE$

```{r}
M <- summary(model1.1)$coefficients
se <- M[2,2]#standard error
m<-mean(DURATION)
llb <-  m- qnorm(0.975)*se
ulb <- m + qnorm(0.975)*se
cib <- c(llb, ulb)
print(paste0("the CI of this model, of DURATION is ",  cib[1]," ",cib[2]))
```


```{r}
M <- summary(model1.3)$coefficients
se <- M[2,2]#standard error
m<-mean(TIME)
llb <-  m- qnorm(0.975)*se
ulb <- m + qnorm(0.975)*se
cib <- c(llb, ulb)
print(paste0("the CI of this model, of TIME is ",  cib[1]," ",cib[2]))
```



```{r}
M <- summary(model1.4)$coefficients
se <- M[2,2]#standard error
m<-mean(LOG.1.DURATION.)
llb <-  m- qnorm(0.975)*se
ulb <- m + qnorm(0.975)*se
cib <- c(llb, ulb)
print(paste0("the CI of this model, of LOG.1.DURATION. is ", cib[1]," ",cib[2]))
```

## Conclusion
If we look at the analysis we can see that early detection is not relevant as there is no clear relationship between morbidity and early detection as seen in the variable DURATION and LOG.1.DURATION, what is closely related is the time in which you have suffered from the disease (TIME).  

## Bibliografy
Slide logistic regression. Jan Graffelman 

medlineplus. https://medlineplus.gov/spanish/lupus.html 

## Apendix
you can find Rmarkdown on this github:
https://github.com/rafaelcastilla/Project_Statistics
